---
title: "HW6_answer"
author: "Guojing Wu | gw2383"
date: "3/31/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = F, 
                      warning = F)
library(tidyverse)
library(nlme)
theme_set(theme_bw())
```

## Problem 1

$$
\begin{split}
varY_{i,j} 
&= var(\mu + b_{i} + e_{i,j}) \\
&= var(b_{i}) + var(e_{i,j}) \\
&= \sigma_{b}^{2} + \sigma_{e}^{2} \\
cov(Y_{i,j},Y_{i,k})
&= E(Y_{i,j}Y_{i,k}) - E(Y_{i,j})E(Y_{i,k}) \\
&= E[(\mu + b_{i} + e_{i,j})(\mu + b_{i} + e_{i,k})] - E(Y_{i,j})E(Y_{i,k}) \\
&= E(\mu^{2} + b_{i}^{2} + 2\mu b_{i} + \mu e_{i,j} + \mu e_{i,k} + b_{i}e_{i,j} + b_{i}e_{i,k} + e_{i,j}e_{i,k}) - \mu^{2} \\
\end{split}
$$

Since $b_{i} \sim N(0, \sigma_{b}^{2})$, $(\frac{b_{i}}{\sigma_{b}})^{2} \sim \chi_{1}^{2}$, so $E(b_{i}^{2}) = \sigma_{b}^{2}$. And since $e_{i,j},e_{i,k}$ are independent, $E(e_{i,j}e_{i,k}) = E(e_{i,j})(e_{i,k})$. So, 

$$
\begin{split}
cov(Y_{i,j},Y_{i,k})
&= \mu^{2} + \sigma_{b}^{2} - \mu^{2} \\
&= \sigma_{b}^{2} \\
corr(Y_{i,j},Y_{i,k})
&= \frac{cov(Y_{i,j},Y_{i,k})}{\sqrt{varY_{i,j}varY_{i,k}}} \\
&= \frac{\sigma_{b}^{2}}{\sigma_{b}^{2} + \sigma_{e}^{2}} \\
\end{split}
$$

This suggest that the covariance pattern is compound symmetry

## Problem 2

```{r}
data_dist = read.table("HW6-dental.txt", header = T) %>% 
  mutate(Child = as.factor(Child), 
         Gender = as.factor(Gender))
```

### 1) spaghetti plot

```{r, dpi = 300}
data_dist %>%
  ggplot(aes(x = Age, y = Distance, group = Child, col = Child)) + 
  geom_line()
```

In general, the older the child is, the longer the distance is.

### 2)

We assume that $Y_{1} \sim Y_{16}$ is male and $Y_{17} \sim Y_{27}$ is female. The covariance within same person is $\sigma_{a}^{2}$, the covariance within the same gender is $\sigma_{b}^{2}$:

$$
\begin{split}
E(Y_{i,j}) &= \beta_{0} + \beta_{1}age_{i,j} \\
\Sigma_{Y_{i}} &= 
\begin{pmatrix} 
  \sigma_{a}^{2} + \sigma_{b}^{2} + \sigma_{e}^{2} & \sigma_{a}^{2} + \sigma_{b}^{2} & ... & \sigma_{a}^{2} + \sigma_{b}^{2} \\
   & \sigma_{a}^{2} + \sigma_{b}^{2} + \sigma_{e}^{2} & ... & \sigma_{a}^{2} + \sigma_{b}^{2} \\
   & & ... \\
   & & ... & \sigma_{a}^{2} + \sigma_{b}^{2} + \sigma_{e}^{2}
\end{pmatrix} \\
\bf{\Sigma} &=
\begin{pmatrix} 
  \Sigma_{Y_{1}} & \sigma_{b}^{2}E & ... & \sigma_{b}^{2}E & 0 & 0 & & 0 \\
   & \Sigma_{Y_{2}} & ... & \sigma_{b}^{2}E & 0 & 0 & & 0 \\
   & & ... & \\
   & & & \Sigma_{Y_{16}} & 0 & 0 &  & 0 \\
   & & & & \Sigma_{Y_{17}} & \sigma_{b}^{2}E & ... & \sigma_{b}^{2}E \\
   & & & & & \Sigma_{Y_{18}} & ... & \sigma_{b}^{2}E \\
   & & & & & & ... & \\
   & & & & & & ... & \Sigma_{Y_{27}} \\
\end{pmatrix}
\end{split}
$$

### 3)

#### i)

```{r}
CompSymm <- gls(Distance ~ Age + Gender, data_dist, correlation = corCompSymm(form = ~ 1 | Child), weights = varIdent(form = ~ 1| Age), method = "REML")

Expo <- gls(Distance ~ Age + Gender, data_dist, correlation = corExp(form = ~ 1 | Child), weights = varIdent(form = ~ 1| Age), method = "REML")

Auto <- gls(Distance ~ Age + Gender, data_dist, correlation = corAR1(form = ~ 1 | Child), weights = varIdent(form = ~ 1| Age), method = "REML")

coeff = rbind(CompSymm$coefficients, Expo$coefficients, Auto$coefficients)
rownames(coeff) = c("CompSymmetry", "Exponential", "Autoregressive")
coeff %>% knitr::kable()
```

By the table above, the coefficient parameter estimates obtained by three methods are very similar.

#### ii)

```{r}
varfunc = rbind(CompSymm$modelStruct$varStruct, Expo$modelStruct$varStruct, Auto$modelStruct$varStruct)
varfunc = cbind(rep(0, nrow(varfunc)), varfunc)
rownames(varfunc) = c("CompSymmetry", "Exponential", "Autoregressive")
colnames(varfunc) = c("8", "10", "12", "14")
varfunc %>% exp() %>% knitr::kable()
```

By the table above, the variance function of different age obtained by three methods are very similar.

#### iii)

```{r}
getVarCov(CompSymm)
getVarCov(Expo)
getVarCov(Auto)
```

By the table above, the covariance matrices matches each pattern. And since it's equal intervals of time, the autoregression and exponential methods are pretty similar.